Robust Clustering for Ad Hoc Cognitive Radio Network

### 2017-12-10 journal-draft/ettdoc_JG.tex
revison based on james' comments on 5th Dec

In this paper, we analytically study the above mentioned trade-off which we term in the following the \textit{CRN robust clustering problem}.
We show it to be an NP-hard problem under certain assumptions, and furthermore study centralized as well as distributed algorithms.
We propose an alternate metric to measure cluster robustness in contrast to previous works~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
We claim that cluster robustness can not be indicated merely by the average number of CCs of a cluster, but by the ability of the cluster to uphold over time despite random primary user activity.
Our proposed distributed scheme extends our previous work ROSS (Robust Spectrum Sharing)~\cite{Li11_ROSS} by additionally incorporating control over the size of a cluster.
Throughout this paper, we call these newly proposed distributed schemes \textit{variants of ROSS}.
%
%
The rest of the paper is organized as follows:
In Section~\ref{related_work}, we review related work in particular with respect to clustering techniques in CRN.
We also discuss in more detail the relation between the contribution in this paper and our previous work in~\cite{Li11_ROSS}.
Our system model as well as the problem statement with respect to the robust clustering problem are presented in Section~\ref{sec:model}. 
The main contribution, the centralized and distributed solutions are introduced in Section~\ref{centralized_solution} and~\ref{ross} respectively.
%The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Extensive performance evaluation is given in Section~\ref{performance} before we conclude our work in Section~\ref{conclusion}.

@@

% XXX Di, these previous three sentences do not make sense: Robustness is actually really only the number of CCs. If you want to add the fact  that the size should be controlled, in particular that not too small clusters are formed, you need to motivate this differently, for example by the sensing accuracy, as I did in the introduction. Please try to reformulate these three sentences accordingly
% XXXX I changed the three sentences, see below.
Robustness of a cluster is its ability to uphold communication among the cluster members despite the influence of the active primary users.
Thus, to achieve better robustness, a clear component of an optimization metric needs to be the amount of CCs among each formed cluster. 
However, this can lead in an extreme situation to a large amount of singleton clusters, if the size of the clusters is not controlled simultaneously. 
A large amount of singleton clusters, as discussed reduced spectrum sensing accuracy through cooperative sensing, as well as being not desirable from different other perspectives.
Thus, we essentially propose to include this trade-off in the optimization process of building clusters, captured in the following definition:

\begin{mydef}
\label{def_centralized_clustering}
For a set of CR nodes $\mathcal{N}$, the \textbf{CRN robust clustering problem} is to determine a set of clusters $\mathcal{T}$, where
\begin{enumerate}
\setlength{\itemindent}{.05in}
\item the intersection of any two clusters in $\mathcal{T}$ results in the empty set;
\item the union of all clusters in $\mathcal{T}$ results in$\mathcal{N}$;
\item the sum over $f(C)$ for all clusters $C$ is maximized, where the number of common channels for cluster $C$ is denoted as $f(C)$;
\item all cluster sizes are within the range $\big[\delta_1, \delta_2\big]$, with $\delta_1, \delta_2\in \mathbb{Z}^+$ and $\delta_1 \leq \delta_2$.
The range can also be expressed as $[\delta-u, \delta+u]$, where $\delta$ is the desired cluster size, $u$ is a constant and $\delta-u > 0$.
\item When the cluster size is larger or smaller than the range $\big[\delta_1, \delta_2\big]$, $f(C)$ is defined as 0, i.e. singleton clusters may be formed but do mot contribute to the objective function.

\end{enumerate}
\end{mydef}
% XXX Di: What does the last statement mean with respect to the size of C being allowed to be 1. Does that mean for all C in T, the size is allowed to be 1? Doesn't that contradict what you have been saying in the definition before, where you come up with the range ?
% XXXX 
%1. we have a function f(C) to denote the number of CCs, if cluster size is in the scope, fine; if not, f(C) equals zero.
%2. it is OK for every CR node to form their own cluster. Then according to rule 3, f(C) is zero and we obtain a 'bad' clustering result.
% XXXX

The decision version of this problem is to determine whether there exists a set of clusters, say $\mathcal{X}$, so that $\cup_{C\in\mathcal{X}} C = \mathcal{N}$, and $\sum_{C\in \mathcal{X}} f(C) \geqslant \lambda$ where $\lambda$ is a positive integer number.
We have the following theorem on the problem's complexity.
\begin{theorem}
\label{theorem1}
The robust clustering problem in CRN is NP-hard, when $\delta_1=2$ and $\delta_2 > 3$.
%when the maximum size of clusters is larger than 3
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is given in Appendix~\ref{proof_theorem1}.

\section{Centralized Solution for Robust Clustering}
\label{centralized_solution}
Given the fact that the complexity of the CRN clustering problem is determined, we consider in this section a centralized solution for the problem.
The motivation is mainly to have later on a comparison scheme for distributed algorithms, that we present in the next section.
Assuming some global knowledge of the CRN to be given at some point in the network, a centralized solution \ie the locations of primary users and their working channels, and the locations of secondary users and available channels on them, we can propose a centralized scheme.
We obtain the set of $\mathcal{S}$ which contains all the clusters in $\mathcal{N}$, \ie $\mathcal{S}=\{C_1, C_2,\ldots,C_i, \ldots, C_{|\mathcal{S}|}\}$ \footnote{The subscript $i$ means the $i$-th cluster in $\mathcal{S}$.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = \mathcal{N}$.
The proposed centralized solution formulates the problem in Definition~\ref{def_centralized_clustering} as an optimization problem which is solved with standard software packages.
The optimization decides on the clusters according to the following optimization formulation, which is a binary linear programming problem and can be solved by many available solvers.

\begin{equation}
\begin{aligned}
     &\max\limits_{y_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (y_i\cdot \frac{q_{ij}}{|C_i|} - y_i\cdot p(C_i)) \\
     &\text{subject to}   && \Sigma_{i=1}^M x_{ij} = 1, \text{for}\,\,\forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|\cdot y_i,\, \text{for}\,\,\forall i=1, \ldots, M \\
  % &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots M\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}
$N$ is the total number of CR users in network $\mathcal{N}$.
$y_i$ and $x_{ij}$ are two binary variables.
Being either 1 or 0, $y_i$ denotes whether the $i$-th cluster $C_i$ in $\mathcal{S}$ is chosen or not.
$x_{ij}$ indicates whether the CR node $j$ resides in the cluster $C_i$, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$.

@@

%The optimization is a binary linear programming problem.
As to the objective function, the sum of the first items is the sum of CCs of the clusters which constitute the CRN.
In the second item, we propose $p(C_i)$ which is a size-related weight for cluster $C_i, i\in{1,\cdots, M}$.
$p(C_i)$ is positively related with the difference between $C_i$'s size and the desired size $\delta$.
The desired size $\delta$ is pre-decided based on the capability of the CR users and the tasks to be conveyed.
When $y_i$ is 1 ($C_i$ is chosen) but $|C_i|$ doesn't equal to the desired cluster size $\delta$, the second item is negative, which contradicts the direction of the optimization.
Thus the second item discourages the appearance the clusters whose sizes deviate from the desired cluster size $\delta$.
The weight $p(C_i)$ with respect to different cluster sizes is as follows,
%$$
%p(C_i) = \left\{ \begin{array}{rl}
%0 &\mbox{ if $|C_i|=\delta$} \\
%\rho_1 &\mbox{if $||C_i|-1|=\delta$} \\
%\rho_2 &\mbox{if $||C_i|-2|=\delta$} \\
%\vdots\\
%\rho_\sigma &\mbox{if $||C_i|-\sigma|=\delta$} \\
%\end{array} \right.
%$$

$$
p(C_i) = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\rho_1 &\mbox{if $|C_i|=\delta \pm 1$ } \\
\rho_2 &\mbox{if $|C_i|=\delta \pm 2$} \\
\vdots\\
\rho_\sigma &\mbox{if $|C_i|=\delta \pm \sigma$} \\
\end{array} \right.
$$
where $ 0 < \rho_1< \rho_2 < \cdots < \rho_\sigma$.
Please note that here we adopt the desired cluster size in stead of the range for cluster sizes which is described in Definition~\ref{problem}.
Because in implementation, we usually need a large range to guarantee a feasible result for the optimization problem.
By adopting the desired size and the punishment for deviating from it, we actually set a large range of cluster sizes and meanwhile a better control of the resulted sizes.

The difficulty of using this method lies in obtaining the set $\mathcal{S}$.
In the worst case, \ie every CR node communicates directly with any other node and the CRN forms a full connected graph, the size of $\mathcal{S}$ is $\Sigma_{r=1}^{N}\ {N \choose r} = 2^N-1$.

@@

In this section we introduce our distributed clustering schemes.
With the variants of ROSS, CR nodes form clusters based on their own available channels, as well as the available channels of the nodes in their neighborhood.
This clarified and conducted through a series of interactions on the control channel.
All variants of ROSS consist of two cascaded phases: \textit{Cluster formation} and \textit{Membership clarification}, as shown in Figure~\ref{flowChartROSS}.
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.9\linewidth]{flow_chart.pdf}
	\caption{Processing steps of ROSS}
	\label{flowChartROSS}
\end{figure}
In the first phase, clusters are initially formed such that every CR user becomes either cluster head or cluster member.
During this phase, size control is already realized, however, memberships might not be efficient with respect to robustness while also not being necessarily unique.
This is addressed in the second phase, where non-overlapping clusters are formed in a way that the CCs of the involved clusters are predominantly increased.

@@

\item \textit{Individual connectivity degree} $d_i$: $d_i=\sum_{j\in \text{Nb}(i)}\vert K_i\cap K_j\vert$. 
$d_i$ is the total number of the CCs between node $i$ and each of its neighbors, it does not reflect the amount of jointly common channels among all neighbors of $i$.
\item \textit{Neighborhood connectivity degree} $g_i$: In contrast, $g_i$ is the number of CCs which are available for $i$ and all of its neighbors, thus
$g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It therefore represents the ability of $i$ to form a robust cluster with its neighbors.
\end{itemize}
Individual connectivity degree $d_i$ and neighborhood connectivity degree $g_i$ together form the \textit{connectivity vector} $(d_i, g_i)$.
The connectivity vector is determined by every secondary user and is then broadcasted.
Figure~\ref{fig1} illustrates the computation of the connectivity vectors for a CRN, where a dashed edge indicates one end node is within the other's transmission range, while the number along the dashed line is the number of common channels between the two end nodes.
In this example, the sets of available channels on the nodes are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, \\K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, \\K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, \\K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. 
Figure~\ref{fig1} shows in particular the resulting connectivity vector per node.
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Illustration of the resulting connectivity vector $(d_i, g_i)$ for each node of an example CRN.}
	\label{fig1}
\end{figure}

\subsubsection{Determining Cluster Heads and Forming Clusters}
Given the connectivity vector per node, the procedure of determining cluster heads is as follows.
Each CR node decides whether it is a cluster head by comparing its connectivity vector with all neighboring connectivity vectors.
When CR node $i$ has lower individual connectivity degree than any of its neighbors except for those which have already been identified as cluster heads, node $i$ becomes a cluster head.
If there is a CR node $j$ in $i$'s neighborhood, which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ while the connectivity degree of $j$ is lower than for all other nodes in its neighborhood (except for nodes that already declared themselves as heads) then out of $i$ and $j$ the node with higher neighborhood connectivity degree will become cluster head.
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes the cluster head.
%
The node which is identified as cluster head broadcasts a message to notify its neighbors of this status update. 
As a consequence, all neighbors - which have not become cluster head themselves -  become cluster members of this cluster head.
In this step, nodes can become member of multiple clusters, depending on how many neighbors declare themselves as cluster heads.
\footnote{The issues arising out of cluster heads in the neighborhood of a newly formed cluster head are addressed in Section~\ref{ross_p1_guarantee_ccc} and \ref{ross_p2_cluster_pruning}}
During the whole phase I, whenever a CR node becomes cluster head, or the cluster composition changes, the cluster head broadcasts new/updated information about the cluster structure, in particular the new/ update sets of available channels regarding itself and all its cluster members.
Pseudo code regarding this process, i.e. the cluster head decision and the initial cluster formation, is also provided as Algorithm~\ref{alg0} in the appendix.

@@

\subsubsection{The Existence of Common Channels}
\label{ross_p1_guarantee_ccc}
After executing Algorithm~\ref{alg0}, several formed clusters may not possess any CCs.
As decreasing the cluster size usually increases CCs within a cluster, the next step is to decrease the the cluster size accordingly.
This is done by following sequence of removing nodes according to an ascending list of nodes regarding their number of common channels between them and the cluster head. 
In other words, the cluster member which has the least common channels with the cluster head will be removed first.
When there are multiple nodes having the same amount of common channels with the cluster head, the node whose elimination results in more common channels will be removed.
In case of a tie, it can be broken by removing the node with smaller node ID.
It is possible that cluster heads remove all their neighbors to obtain CCs, which results in a singleton cluster.
The pseudo code for this procedure is given as Algorithm~\ref{alg_size_control_available_CCC}.
As for the nodes which are removed from a cluster, they restore their original individual connectivity degrees, then execute Algorithm~\ref{alg0} and become either cluster heads or get included into other clusters, see also Theorem~\ref{clustering:theorem}.

@@

\begin{itemize}
	\item $\mathcal{P}$, the set of players in the game, which are the debatable nodes in our problem.
	\item $\mathcal{R} = \cup S_i, i\in \mathcal{P}$, the set of the resources for players to choose. In our problem, $S_i$ is the set of the claiming clusters of $i$, and $\mathcal{R}$ is the set of all claiming clusters.
	\item Strategy space $\sum_i, i \in \mathcal{P}$, $\sum_i$ is the set of the claiming clusters $S_i$.
	As debatable node $i$ is supposed to choose only one claiming cluster, only a single resource will be allocated to $i$.%, accordingly this congestion game is a singleton game.
	\item 	The cost function $f(C)$ regarding resource $C$. 
	$f(C) = \Delta\vert K^i(C)|, C\in S_i$, which represents the decreased number of CCs in cluster $C$ when debatable node $i$ joins $C$.
	As to cluster $C\in S_i$, the decrease of CCs caused by accepting the debatable nodes is $\sum_{i:C\in S_i, i\rightarrow C} \Delta\vert K^i(C) \vert$. 
$i\rightarrow C$ means $i$ joins cluster $C$.
Obviously this function is non-decreasing with respect to the number of nodes joining cluster $C$.

@@

%This is the global objective we want to minimize.
\end{itemize}

@@

\section{Performance Evaluation}
\label{performance}
Taking the final clustering result of ROSS into account for our toy example shown in Figure~\ref{final_clustering_ross}, we can compare the outcome with our centralized scheme proposed in Equation~\ref{centralized_opt} as well as the state-of-the-art algorithm SOC~\cite{LIU_TMC11_2}.
Those corresponding results of the later two schemes are shown in Figure~\ref{fig:final_clustering}.
We observe for this example case that ROSS and the centralized scheme achieve cluster sizes that are more balanced, while SOC leads to a larger variance in terms of the cluster size.
Regarding the amount of CCs, the same observation holds. 
\begin{figure}[ht]
\begin{center}
\subfigure[Generated by SOC]{\includegraphics[width=0.435\linewidth]{final_clustering_soc}\label{fig:final_clustering_soc}}
\hspace{0.15 in}
\subfigure[Generated by the centralized clustering scheme]{\includegraphics[width=0.435\linewidth]{final_clustering_LP}\label{fig:final_clustering_LP}}
\end{center}
\caption{Final clusters formed by SOC as well as the centralized clustering scheme.}
\label{fig:final_clustering}
\end{figure}

In the following, we are interested in a more general performance comparison regarding clustering in CRNs.
We therefore present in the following an extensive evaluation study.
We base our evaluations on simulations, and consider the following comparison schemes:
\begin{itemize}
\item ROSS without size control: ROSS-DGA, ROSS-DFA;
\item ROSS with size control, \ie ROSS-$\delta$-DGA and ROSS-$\delta$-DFA where $\delta$ is the desired cluster size;
\item SOC~\cite{LIU_TMC11_2}, a distributed clustering scheme pursuing cluster robustness;
\item Centralized robust clustering scheme; in our evaluations we use the built-in function $bintprog$ of MATLAB to solve the corresponding integer optimization problem given in Equation~\ref{centralized_opt}
\end{itemize}

>>>>

In this paper, we analytically study the above mentioned trade-off which we term in the following the \textit{CRN robust clustering problem}.
We show it to be an NP-hard problem under certain assumptions, and furthermore study centralized as well as distributed algorithms.
We propose an alternate metric to measure cluster robustness in contrast to previous works~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
We claim that cluster robustness can not be indicated merely by the average number of CCs of a cluster, but by the ability of the cluster to uphold over time despite random primary user activity.
Our proposed distributed scheme extends our previous work ROSS (Robust Spectrum Sharing)~\cite{Li11_ROSS} by additionally incorporating control over the size of a cluster.
Throughout this paper, we call these newly proposed distributed schemes \textit{variants of ROSS}.
%

The rest of the paper is organized as follows:
In Section~\ref{related_work}, we review related work in particular with respect to clustering techniques in CRN.
We also discuss in more detail the relation between the contribution in this paper and our previous work in~\cite{Li11_ROSS}.
Our system model as well as the problem statement with respect to the robust clustering problem are presented in Section~\ref{sec:model}. 
The main contribution, the centralized and distributed solutions are introduced in Section~\ref{centralized_solution} and~\ref{ross} respectively.
%The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Extensive performance evaluation is given in Section~\ref{performance} before we conclude our work in Section~\ref{conclusion}.

@@

% XXX Di, these previous three sentences do not make sense: Robustness is actually really only the number of CCs. If you want to add the fact  that the size should be controlled, in particular that not too small clusters are formed, you need to motivate this differently, for example by the sensing accuracy, as I did in the introduction. Please try to reformulate these three sentences accordingly
% XXXX I changed the three sentences, see below.
Robustness of a cluster is its ability to uphold communication among the cluster members despite the influence of the active primary users.
Thus, to achieve better robustness, a clear component of an optimization metric needs to be the amount of CCs among each formed cluster. 
However, this can lead in an extreme situation to a large amount of singleton clusters, if the size of the clusters is not controlled simultaneously. 
As discussed, a large amount of singleton clusters reduces spectrum sensing accuracy through cooperative sensing, as well as being not desirable from different other perspectives (routing, coordination with respect to channel vacation).
Thus, we essentially propose to include this trade-off in the optimization process of building clusters, captured in the following definition:

\begin{mydef}
\label{def_centralized_clustering}
For a set of CR nodes $\mathcal{N}$, the \textbf{CRN robust clustering problem} is to determine a set of clusters $\mathcal{T}$, where:
\begin{enumerate}
\setlength{\itemindent}{.05in}
\item The intersection of any two clusters in $\mathcal{T}$ results in the empty set.
\item The union of all clusters in $\mathcal{T}$ results in $\mathcal{N}$.
\item For all clusters with size within the range $\big[\delta_1, \delta_2\big]$, with $\delta_1, \delta_2\in \mathbb{Z}^+$ and $\delta_1 \leq \delta_2$, the number of CCs per cluster is $f(C) = K(C)$, i.e. it is given as the number of jointly available CCs for all nodes of that cluster. 
The desired size $\delta$ is within $\big[\delta_1, \delta_2\big]$, which is pre-decided based on the capability of the CR users and the tasks to be conveyed.
%In particular, the range can be expressed as $[\delta-u, \delta+u]$, where $\delta_1= \delta -u, \delta_2=\delta+u$, $\delta$ is referred to as \textbf{desired cluster size}, $u$ is a constant and $\delta-u > 0$.
\item When the cluster size is larger or smaller than the range $\big[\delta_1, \delta_2\big]$, $f(C)$ is defined as 0, i.e. singleton clusters may be formed but does not contribute to the objective function.
\item The sum over $f(C)$ for all clusters $C \in \mathcal{T}$ is maximal.
\end{enumerate}
\end{mydef}
Note that in the above definition, we distinguish between the real number of CCs per cluster, which is given as $K(C)$, and the contribution of each formed cluster towards the objective function, which is given by function $f(C)$.
If the cluster size is within range, the two parameters are the same, but otherwise the distinction is enforce a cost for building clusters that are out of range.
% XXX Di: What does the last statement mean with respect to the size of C being allowed to be 1. Does that mean for all C in T, the size is allowed to be 1? Doesn't that contradict what you have been saying in the definition before, where you come up with the range ?
% XXXX 
%1. we have a function f(C) to denote the number of CCs, if cluster size is in the scope, fine; if not, f(C) equals zero.
%2. it is OK for every CR node to form their own cluster. Then according to rule 3, f(C) is zero and we obtain a 'bad' clustering result.
% XXXX

The decision version of this problem is to determine whether there exists a set of clusters, say $\mathcal{T}$, so that $\cup_{C\in\mathcal{T}} C = \mathcal{N}$, and $\sum_{C\in \mathcal{T}} f(C) \geqslant \lambda$ where $\lambda$ is a positive integer number.
We have the following theorem on the problem's complexity.
\begin{theorem}
\label{theorem1}
The robust clustering problem in CRN is NP-hard, when $\delta_1=2$ and $\delta_2 > 3$.
%when the maximum size of clusters is larger than 3
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is given in Appendix~\ref{proof_theorem1}.

\section{Centralized Solution for Robust Clustering}
\label{centralized_solution}
%XXXXXX
%Di: The first paragraph has been revised based on the first batch of comments for the camera-ready version, but somehow get lost in the last version. I take some parts from your revision and keep some from my revision.
We now turn to algorithms that can solve the CRN robust clustering problem, despite its complexity. 
We initially consider a centralized solution for this problem.
Assuming some global knowledge of the CRN to be given at some point in the network, \ie the locations of primary users and their working channels, and the locations of secondary users and available channels on them, we can propose a centralized scheme.
We obtain the set of $\mathcal{S}$ which contains all the clusters which satisfy the definition of cluster in Section~\ref{sec:model}.
$\mathcal{S}$ is basically a powerset of $\mathcal{N}$, which nevertheless is restricted by connectivity and spectrum availability in the network.
There are $\mathcal{S}=\{C_1, C_2, \ldots, C_{|\mathcal{S}|}\}$ \footnote{The subscript $i$ means the $i$-th cluster in $\mathcal{S}$.} and $|\mathcal{S}| = M$.
Then the problem as defined in Definition~\ref{def_centralized_clustering} can be formulated as a binary linear programming problem and can be solved by many available solvers:
\begin{equation}
\begin{aligned}
     &\max\limits_{y_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (y_i\cdot \frac{q_{ij}}{|C_i|} - y_i\cdot p(C_i)) \\
     &\text{subject to}   && \Sigma_{i=1}^M x_{ij} = 1, \text{for}\,\,\forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|\cdot y_i,\, \text{for}\,\,\forall i=1, \ldots, M \\
  % &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots M\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}
$y_i$ and $x_{ij}$ are the two sets of binary optimization variables.
Being either 1 or 0, $y_i$ denotes whether the $i$-th cluster $C_i$ in $\mathcal{S}$ is chosen or not.
$x_{ij}$ indicates whether the CR node $j$ resides in the cluster $C_i$, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$.

@@

%The optimization is a binary linear programming problem.
As to the objective function, the sum of the first items is the sum of CCs of the clusters which constitute the CRN.
In the second item, we propose $p(C_i)$ which is a size-related weight for cluster $C_i, i\in{1,\cdots, M}$.
$p(C_i)$ is positively related with the difference between $C_i$'s size and the desired size $\delta$.
When $y_i$ is 1 ($C_i$ is chosen) but $|C_i|$ doesn't equal to the desired cluster size $\delta$, the second item is negative, which contradicts the direction of the optimization.
Thus the second item discourages the appearance the clusters whose sizes deviate from the desired cluster size $\delta$.
The weight $p(C_i)$ with respect to different cluster sizes is as follows,
%$$
%p(C_i) = \left\{ \begin{array}{rl}
%0 &\mbox{ if $|C_i|=\delta$} \\
%\rho_1 &\mbox{if $||C_i|-1|=\delta$} \\
%\rho_2 &\mbox{if $||C_i|-2|=\delta$} \\
%\vdots\\
%\rho_\sigma &\mbox{if $||C_i|-\sigma|=\delta$} \\
%\end{array} \right.
%$$

$$
p(C_i) = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\rho_1 &\mbox{if $|C_i|=\delta \pm 1$ } \\
\rho_2 &\mbox{if $|C_i|=\delta \pm 2$} \\
\vdots\\
\rho_\sigma &\mbox{if $|C_i|=\delta \pm \sigma$} \\
\end{array} \right.
$$
where $ 0 < \rho_1< \rho_2 < \cdots < \rho_\sigma$.
Note that here we adopt the desired cluster size in stead of the range for cluster sizes which is described in Definition~\ref{problem}.
Because in implementation, we usually need a large range to guarantee a feasible result for the optimization problem.
By adopting the desired size and the punishment for deviating from it, we actually set a large range of cluster sizes and meanwhile a better control of the resulted sizes.

The difficulty of using this method lies in obtaining the set $\mathcal{S}$.
In the worst case, \ie if every CR node can communicate directly with any other node, then the CRN forms a full connected graph and therefore the size of the powerset $\mathcal{S}$ is $\Sigma_{r=1}^{N}\ {N \choose r} = 2^N-1$.

@@

In this section we introduce our distributed clustering schemes.
With the variants of ROSS, CR nodes form clusters based on their own available channels, as well as the available channels of the nodes in their neighborhood.
This is clarified and conducted through a series of interactions on the control channel.
All variants of ROSS consist of two cascaded phases: \textit{Cluster formation} and \textit{Membership clarification}, as shown in Figure~\ref{flowChartROSS}.
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.9\linewidth]{flow_chart.pdf}
	\caption{Processing steps of ROSS}
	\label{flowChartROSS}
\end{figure}
In the first phase, clusters are initially formed such that every CR user becomes either cluster head or cluster member.
During this phase, size control is already realized, however, memberships might not be efficient with respect to robustness while also not being necessarily unique.
This is addressed in the second phase, where non-overlapping clusters are formed in a way that the CCs of the involved clusters are predominantly increased.

@@

\item \textit{Individual connectivity degree} $d_i$: $d_i=\sum_{j\in \text{Nb}(i)}\vert K_i\cap K_j\vert$. 
$d_i$ is the total number of \textit{pairwise} CCs between node $i$ and each of its neighbors. Be aware that it does not reflect the amount of \textit{jointly} common channels among all neighbors of $i$.
\item \textit{Neighborhood connectivity degree} $g_i$: In contrast, $g_i$ is the number of CCs which are \textit{jointly} available for $i$ and all of its neighbors, thus
$g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It therefore represents the ability of $i$ to form a robust cluster with its neighbors.
\end{itemize}
Individual connectivity degree $d_i$ and neighborhood connectivity degree $g_i$ together form the \textit{connectivity vector} $(d_i, g_i)$.
The connectivity vector is determined by every secondary user and is then broadcasted.
Figure~\ref{fig1} illustrates the computation of the connectivity vectors for a CRN, where a dashed edge indicates one end node is within the other's transmission range, while the number along the dashed line is the number of common channels between the two end nodes.
In this example, the sets of available channels on the nodes are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, \\K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, \\K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, \\K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. 
Figure~\ref{fig1} shows in particular the resulting connectivity vector per node.
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Illustration of the resulting connectivity vector $(d_i, g_i)$ for each node of an example CRN.}
	\label{fig1}
\end{figure}

\subsubsection{Determining Cluster Heads and Forming Clusters}
Given the connectivity vector per node, the procedure of determining cluster heads is as follows.
Each CR node decides whether it is a cluster head by comparing its connectivity vector with all neighboring connectivity vectors.
When CR node $i$ has lower individual connectivity degree than any of its neighbors except for those which have already been identified as cluster heads, node $i$ becomes a cluster head.
If there is a CR node $j$ in $i$'s neighborhood, which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ while the connectivity degree of $j$ is lower than for all other nodes in its neighborhood (except for nodes that already declared themselves as heads) then out of $i$ and $j$ the node with higher neighborhood connectivity degree will become cluster head.
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes the cluster head.
%
The node which is identified as cluster head broadcasts a message to notify its neighbors of this status update. 
As a consequence, all neighbors - which have not become cluster head themselves -  become cluster members of this cluster head.
In this step, nodes can become member of multiple clusters, depending on how many neighbors declare themselves as cluster heads.
\footnote{The issues arising out of cluster heads in the neighborhood of a newly formed cluster head are addressed in Section~\ref{ross_p1_guarantee_ccc} and \ref{ross_p2_cluster_pruning}}
During the whole phase I, whenever a CR node becomes cluster head, or the cluster composition changes, the cluster head broadcasts new/updated information about the cluster structure, in particular the new/updated sets of available channels regarding itself and all its cluster members.
Pseudo code regarding this process, i.e. the cluster head decision and the initial cluster formation, is provided as Algorithm~\ref{alg0} in the appendix.

@@

\subsubsection{The Existence of Common Channels}
\label{ross_p1_guarantee_ccc}
After executing Algorithm~\ref{alg0}, several formed clusters may not possess any CCs.
As decreasing the cluster size usually increases CCs within a cluster, the next step is to decrease the cluster size accordingly.
This is done by the following sequence of removing nodes according to an ascending list of nodes regarding their number of common channels between them and the cluster head. 
In other words, the cluster member which has the least common channels with the cluster head will be removed first.
When there are multiple nodes having the same amount of common channels with the cluster head, the node whose elimination results in more common channels will be removed.
In case of a tie, it can be broken by removing the node with smaller node ID.
It is possible that cluster heads remove all their neighbors to obtain CCs, which results in a singleton cluster.
The pseudo code for this procedure is given as Algorithm~\ref{alg_size_control_available_CCC}.
As for the nodes which are removed from a cluster, they restore their original individual connectivity degrees, then execute Algorithm~\ref{alg0} and become either cluster heads or get included into other clusters, see also Theorem~\ref{clustering:theorem}.

@@

\begin{itemize}
	\item $\mathcal{P}$, the set of players in the game, which are the debatable nodes in our problem.
	\item $\mathcal{R} = \cup S_i, i\in \mathcal{P}$, the set of the resources for players to choose. In our problem, $S_i$ is the set of the claiming clusters of $i$, and $\mathcal{R}$ is the set of all claiming clusters.
	\item Strategy space $\sum_i, i \in \mathcal{P}$, $\sum_i$ is the set of the claiming clusters $S_i$.
	As debatable node $i$ is supposed to choose only one claiming cluster, only a single resource will be allocated to $i$.%, accordingly this congestion game is a singleton game.
	\item 	The cost function $f(C)$ regarding resource $C$. 
	$f(C) = \Delta\vert K^i(C)|, C\in S_i$, which represents the decreased number of CCs in cluster $C$ when debatable node $i$ joins $C$.
	As to cluster $C\in S_i$, the decrease of CCs caused by accepting the debatable nodes is $\sum_{i:C\in S_i, i\rightarrow C} \Delta\vert K^i(C) \vert$. 
$i\rightarrow C$ means $i$ joins cluster $C$.
Obviously this function is non-decreasing with respect to the number of nodes joining cluster $C$.
\end{itemize}

@@

%This is the global objective we want to minimize.

@@

\section{Performance Evaluation}
\label{performance}
Taking the final clustering result of ROSS into account for our toy example shown in Figure~\ref{final_clustering_ross}, we can compare the outcome with our centralized scheme proposed in Equation~\ref{centralized_opt} as well as the state-of-the-art algorithm SOC~\cite{LIU_TMC11_2}.
Those corresponding results of the latter two schemes are shown in Figure~\ref{fig:final_clustering}.
We observe for this example case that ROSS and the centralized scheme achieve cluster sizes that are more balanced, while SOC leads to a larger variance in terms of the cluster size.
Regarding the amount of CCs, the same observation holds. 
\begin{figure}[ht]
\begin{center}
\subfigure[Generated by SOC]{\includegraphics[width=0.435\linewidth]{final_clustering_soc}\label{fig:final_clustering_soc}}
\hspace{0.15 in}
\subfigure[Generated by the centralized clustering scheme]{\includegraphics[width=0.435\linewidth]{final_clustering_LP}\label{fig:final_clustering_LP}}
\end{center}
\caption{Final clusters formed by SOC as well as the centralized clustering scheme.}
\label{fig:final_clustering}
\end{figure}

In the following, we are interested in a more general performance comparison regarding clustering in CRNs.
We therefore present an extensive evaluation study.
We base our evaluations on simulations, and consider the following comparison schemes:
\begin{itemize}
\item ROSS without size control: ROSS-DGA, ROSS-DFA;
\item ROSS with size control, \ie ROSS-$\delta$-DGA and ROSS-$\delta$-DFA where $\delta$ is the desired cluster size;
\item SOC~\cite{LIU_TMC11_2}, a distributed clustering scheme pursuing cluster robustness;
\item Centralized robust clustering scheme; in our evaluations we use the built-in function $bintprog$ of MATLAB to solve the corresponding integer optimization problem given in Equation~\ref{centralized_opt}
\end{itemize}
